Realizing temporal graphs from fastest travel times

### 2023-04-24 MFCS23/MFCS23-shortversion.tex
MFCS short version

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.

@@

    \emph{Connector Gadget.} Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by fast paths, which will make arguing about possible labelings in \textsc{yes}-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and two connector gadgets for each validation gadget.

@@

Furthermore, we use \emph{connector gadgets}, two for each edge selection gadget, and two for every verification gadget. They consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$ and, intuitively, are used to connect many vertex pairs by fast paths, which will make arguing about possible labelings in \textsc{yes}-instances much easier. Finally, we have an \emph{alignment gadget}, which is a star with a center vertex $w^\star$ and a leaf for every other gadget. Intuitively, this gadget is used to relate labels of different gadgets to each other. A formal description of these two gadgets is given in the Appendix.

@@

    \begin{definition}[Travel delays]
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    %Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.
    From the above definition and the definition of the duration of the temporal path $P$ we get the following results.
    \begin{observation}[$\star$]\label{obs:durationPwithWaitingTimes}
        Let $P = (v_0, v_1, \dots, v_p)$ be the underlying path of the temporal path
        $(P, \lambda)$ from $v_0$ to $v_p$.
        Then $d(P,\lambda) = \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1 $.
    \end{observation}
    \begin{observation}[$\star$]\label{obs:travel-delays-both-directions}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of problem \deltaExact.
    For any two incident edges $e_1 = uv$ and $e_2 = vz$ on vertices $u,v,z \in V$, with $e_1 \cap e_2 = v$, we have $\tau_v^{zu} = \Delta - \tau_v^{uz} \pmod \Delta$.
    \end{observation}
\begin{lemma} [$\star$]\label{lemma:unique-fastest-path-allDelays}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of problem \deltaExact,
    and let $(P,\lambda)=(P_{1,k},\lambda)$ be a fastest temporal path from $u=v_1$ to $v=v_k$ on $k$ vertices $v_1,v_2,\dots,v_k$.
    Let us denote with $(P_{1,i}, \lambda)$ the sub-path of temporal path $(P_{1,k},\lambda)$, that starts at $v_1$ and finishes at $v_i$.
    Suppose that for any $v_i$ in $P$ it follows that $(P_{1,i}, \lambda)$ is also the fastest temporal path from $u=v_1$ to $v_i$.
    Then we can determine travel delays on vertices of $P$ using the following equation
    \begin{equation}\label{eq:trav-delays-paths}
        \tau_{v_i}^{v_{i-1},v_{i+1}} = D_{1,i+1} - D_{1,i}, \text{ for all $i \in \{2,3, \dots, k-1\}$.}
    \end{equation}
\end{lemma}
%All of the above is used in the design of the following polynomial-time algorithms.
We are now ready to provide a polynomial-time algorithm for \deltaExact\ when the underlying graph is a tree.
Let $D$ be a matrix from \deltaExact\ and let the underlying graph $G$ of $D$ be a tree on $n$ vertices $\{v_1, v_2, \dots, v_n\}$.
Let $v_i,v_j$ be two arbitrary vertices in $G$. Then we know that there exists a unique (static) path $P$ among them.
Consequently, it follows that the temporal paths $(P_{i,j},\lambda)$ from $v_i$ to $v_j$ and $(P_{j,i},\lambda)$ from $v_j$ to $v_i$ are also unique, up to modulo of the period $\Delta$ of the labeling $\lambda$,
and therefore are the fastest.
Then $D$ is of the following form:
\begin{equation*}
    D_{i,j} =
    \begin{cases}
    0 & \text{if $i = j$}, \\
    1 & \text{if $v_i$ and $v_j$ are neighbours in G}\\
    d(P_{i,j},\lambda) & \text{else}
    \end{cases},
\end{equation*}
where $d(P_{i,j},\lambda)$ is the duration of the (unique) temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$.

@@

\begin{proof}
    Let $D$ be an input matrix for problem \deltaExact\ of dimension $n \times n$.
    In polynomial time we construct the underlying graph $G$ of $D$ and check if it is a tree.
    If at this point we encounter a problem, our algorithm stops and returns a negative answer.

    Let $V(G) = \{v_1, v_2, \dots, v_n\}$.
    We pick an arbitrary vertex $v_i \in V(G)$ and check which vertex $v_j \in V(G)$ is furthest away from it, \ie we find a maximum element in the $i$-th row of the matrix $D$.
    We now take the unique path $P_{i,j}$ in $G$, 
    which has to also be the underlying path of the fastest temporal path from $v_i$ to $v_j$, and using \cref{obs:travel-delays-in-trees,lemma:unique-fastest-path-allDelays} calculate waiting times at all inner vertices.
    We save those values in a matrix $T$, of size $n \times n \times n$, and mark vertices of the path  $P_{i,j}$ as visited.
    Matrix $T$ stores at the position $(k,j,\ell)$ the value corresponding to the travel delay at vertex $v_k$ when traveling from $v_{j}$ to $v_{\ell}$, 
    \ie it stores the value $\tau_{v_k}^{v_{j}, v_\ell}$, where $v_j,v_\ell \in N(v_k)$. All other values of $T$ are set to \textsc{Null}.
    We repeat the above procedure from vertex $v_i$, for vertices that are not marked as visited yet, \ie vertices in $V \setminus P_{i,j}$.
    When we have exhausted the $i$-th row of $D$,
    \ie vertex $v_i$ now reaches every vertex of $G$,
    we continue and repeat the procedure for all other vertices.
    If at any point we get two different values for the same travel delay at a specific vertex, then we stop with the algorithm and return the negative answer.
    If the above procedure finishes successfully we get the matrix $T$ with travel delays for all vertices in $G$, of degree at least $2$.
    The above calculation is performed in polynomial time, as for every vertex $v_i$ we inspect the whole graph once.
    \begin{claim}[$\star$]\label{lemma:matrixT-travelDelays-tree}
        Matrix $T$ consists of travel delays for all vertices of degree at least $2$ in $G$.
    \end{claim}
    Now, given the matrix of travel delays $T$, we can find a labeling $\lambda$ that satisfies $D$.
    We start by fixing the label of an arbitrary edge as $a$, where $a \in [\Delta]$.
    Knowing the label of one edge, and all waiting times in $T$, we can uniquely determine the labels of all other edges.
    More specifically, if we know that $\lambda(v_i v_j) = a$, then for all $v_k \in N(v_i)$ (resp.~$v_{k'} \in N(v_j)$)
    the value $\lambda(v_iv_k) \equiv a + \tau_{v_i}^{v_j,v_k} \pmod \Delta $ (resp.~$\lambda(v_j v_{k'}) \equiv a + \tau_{v_j}^{v_i,v_{k'}} \pmod \Delta $).
    Since there are $\Delta$ options to fix the first label, we can find $\Delta$ different labelings satisfying $D$.
    %Note, w.l.o.g. we can start determining the labeling $\lambda$ by setting $\lambda(v_1v_2) = 1$.
    It is not hard to see, that also the calculation of the labeling $\lambda$ takes polynomial time, as we have to traverse the graph exactly once, to successfully fix the labeling. Therefore, all together the whole algorithm is performed in polynomial time.
\end{proof}

@@

In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Note that one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$, \ie $F = E(G) \setminus E(T)$.

@@

To overcome the above mentioned difficulties, we propose the following high-level strategy.
\begin{itemize}
    \item We identify a small number of ``interesting vertices'' between which we guess representative fastest paths.
    \item We show that we can determine most labels along each path up to an additive constant. We show that the number of labels we cannot determine along such a path is small. 
    \item For each additive constant and each non-determined label along such a path we introduce a variable for an integer linear program (ILP). 
    \item We use constraints in the ILP to ensure that alternative connections (not using the guessed fastest path) between vertex pairs are not too fast.
    \item We compute a solution to the ILP and construct a labeling from that solution. Since the ILP has few variables and not too many constraints, this computation can be done efficiently using Lenstra's Theorem~\cite{Lenstra1983Integer}.
\end{itemize}

For the remainder of this section, we fix the following notation. 
Let $D$ be the input matrix of \deltaExact\ \ie
the matrix of the fastest temporal paths among $n$ vertices, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
%With $F$ we denote a minimum feedback edge set of $G$, and with $k$ the feedback edge number of $G$.
We are now ready to present our FPT algorithm. For an easier readability we split the description and analysis of the algorithm in five subsections.
We start with a preprocessing procedure for graph $G$, where we define a set of interesting vertices which then allows us to guess the desired structures.
Next we introduce some extra properties of our problem, that we then use to create ILP instances and their constraints.
At the end we present how to solve all instances and produce the desired labeling $\lambda$ of $G$, if possible.

@@

%%%%% HM: the two sentences below seem trivial to me.
%Note that we can easily solve \deltaExact\ by guessing the structure of temporal paths among all pairs of vertices.
%But, this would create $f(n,m)$ guesses (where $f(n,m)$ is an exponential function in $|V|=n$ and $|E| = m$), and the running time of the algorithm would be too big.
We approach the problem using some special properties, which allow us to create $f(k)$ different guesses for certain fastest temporal paths (where $f$ will be a double exponential function in the feedback edge number $k$).
We solve each guess by creating an ILP instance with $O(k^3)$ variables and $O(n^c)$ constraints, for some $c \in \mathbb{N}$.
If the ILP instance is a \textsc{yes}-instance then we will show that also $D$ is a \textsc{yes}-instance of \deltaExact.
Moreover, from a positive solution of the ILP instance we then construct the desired labeling $\lambda$.
If all constructed ILP instances are \textsc{no}-instances, then we will show that $D$ is a \textsc{no}-instance of \deltaExact.
Due to Lenstra's Theorem~\cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables.
Therefore, we can solve each of our ILP instances in FPT time with respect to $k$.
\end{comment}
%

\subparagraph{Preprocessing of the input.}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$,
and denote with 
\[
Z = V(G) \setminus V(G').
\]
Then we determine a minimum feedback edge set $F$ of $G'$, and denote with $k = |F|$.
%by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
%Note that $F$ is also a minimum feedback edge set of $G$.
%Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$ and
let $V_1 \subset V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subset V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with 
\[U = V_1 \cup V_2 \cup V_3\]
the set of \emph{vertices of interest}. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie 
\[U^* = \{v \in V(G') \setminus U  :  u \in U, v \in N(u)\}.\]
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, therefore $|U^*| \in O(k^2)$.
From the construction of $Z$ (by iteratively removing vertices of degree one from $G$)
it follows that $Z$ consists of disjoint trees $T_1, T_2, \dots$.
For a tree $T_i$ we denote with $u_i$ the vertex in $G'$ that is a neighbor of a vertex in $T_i$, 
and call it a \emph{clip vertex of the tree $T_i$}.
It follows that there can be many different trees $T_i$ that are incident to the same clip vertex $u_i \in V(G')$,
but each tree $T_i$ is incident to exactly one clip vertex $u_i \in V(G')$.
Since $u_i$ is the only vertex connecting all of the trees $T_i$ incident to it,
from now on we assume that a tree $T_{u_i}$ in $Z$ is a union of trees on vertices from $V(G) \setminus V(G')$,
that are clipped at the same vertex $u_i \in V(G')$.
For each of the trees $T_{u_i}$ in $Z$, we select one vertex $r_i$, that is a neighbor of the clip vertex $u_i$,
and call it \emph{a representative vertex of the tree $T_{u_i}$}.
We now define as $Z^*$ the set of representatives $r_i$ of trees $T_i \in Z$,
where the clip vertex $v_i$ of $T_i$ is a vertex of interest, \ie
\[
Z^* = \{ r_i  :  r_i \in T_i, \text{ where $T_i \in Z$, the clip vertex $u_i$ of $T_i$ is in $U$, and $r_i u_i \in E(G)$}\}.
\]
Since there are $O(k)$ vertices of interest, we get that $|Z^*| \in O(k)$.
Note that determining sets $U$, $U^*$, and $Z^*$ takes linear time.

>>>>

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda,\Delta)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.

@@

    \emph{Connector Gadget.} Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by fast paths, which will make arguing about possible labelings in \textsc{Yes}-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and two connector gadgets for each validation gadget.

@@

Furthermore, we use \emph{connector gadgets}, two for each edge selection gadget, and two for every verification gadget. They consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$ and, intuitively, are used to connect many vertex pairs by fast paths, which will make arguing about possible labelings in \textsc{Yes}-instances much easier. Finally, we have an \emph{alignment gadget}, which is a star with a center vertex $w^\star$ and a leaf for every other gadget. Intuitively, this gadget is used to relate labels of different gadgets to each other. A formal description of these two gadgets is given in the Appendix.

@@

    \begin{definition}[Travel delays]
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    %\begin{equation}\label{eq:def-VertexWaitingTime}
       $\tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.$
    %\end{equation}
    %Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.
    From the above definition and the definition of the duration of the temporal path $P$ we get the following results.
    \begin{observation}[$\star$]\label{obs:durationPwithWaitingTimes}
        Let $P = (v_0, v_1, \dots, v_p)$ be the underlying path of the temporal path
        $(P, \lambda)$ from $v_0$ to $v_p$.
        Then $d(P,\lambda) = \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1 $.
    \end{observation}
    \begin{observation}[$\star$]\label{obs:travel-delays-both-directions}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of problem \deltaExact.
    For any two incident edges $e_1 = uv$ and $e_2 = vz$ on vertices $u,v,z \in V$, with $e_1 \cap e_2 = v$, we have $\tau_v^{zu} = \Delta - \tau_v^{uz} \pmod \Delta$.
    \end{observation}
\begin{lemma} [$\star$]\label{lemma:unique-fastest-path-allDelays}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of problem \deltaExact,
    and let $(P,\lambda)=(P_{1,k},\lambda)$ be a fastest temporal path from $u=v_1$ to $v=v_k$ on $k$ vertices $v_1,v_2,\dots,v_k$.
    Let us denote with $(P_{1,i}, \lambda)$ the sub-path of temporal path $(P_{1,k},\lambda)$, that starts at $v_1$ and finishes at $v_i$.
    Suppose that for any $v_i$ in $P$ it follows that $(P_{1,i}, \lambda)$ is also the fastest temporal path from $u=v_1$ to $v_i$.
    Then we can determine travel delays on vertices of $P$ using the following equation
    %\begin{equation}\label{eq:trav-delays-paths}
        $\tau_{v_i}^{v_{i-1},v_{i+1}} = D_{1,i+1} - D_{1,i},$ for all $i \in \{2,3, \dots, k-1\}$.
    %\end{equation}
\end{lemma}
%All of the above is used in the design of the following polynomial-time algorithms.
We are now ready to provide a polynomial-time algorithm for \deltaExact\ when the underlying graph is a tree.
Let $D$ be a matrix from \deltaExact\ and let the underlying graph $G$ of $D$ be a tree on $n$ vertices $\{v_1, v_2, \dots, v_n\}$.
\begin{comment}
Let $v_i,v_j$ be two arbitrary vertices in $G$. Then we know that there exists a unique (static) path $P$ among them.
Consequently, it follows that the temporal paths $(P_{i,j},\lambda)$ from $v_i$ to $v_j$ and $(P_{j,i},\lambda)$ from $v_j$ to $v_i$ are also unique, up to modulo of the period $\Delta$ of the labeling $\lambda$,
and therefore are the fastest.
Then $D$ is of the following form:
\begin{equation*}
    D_{i,j} =
    \begin{cases}
    0 & \text{if $i = j$}, \\
    1 & \text{if $v_i$ and $v_j$ are neighbours in G}\\
    d(P_{i,j},\lambda) & \text{else}
    \end{cases},
\end{equation*}
where $d(P_{i,j},\lambda)$ is the duration of the (unique) temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$.
\end{comment}

@@

\begin{proof}
    Let $D$ be an input matrix of \deltaExact\ with dimension $n \times n$.
    In polynomial time we construct the underlying graph $G$ of $D$ and check if it is a tree.
    %If at this point we encounter a problem, our algorithm stops and returns a negative answer.

    Let $V(G) = \{v_1, v_2, \dots, v_n\}$.
    We pick an arbitrary vertex $v_i \in V(G)$ and check which vertex $v_j \in V(G)$ is furthest away from it, \ie we find a maximum element in the $i$-th row of the matrix $D$.
    We now take the unique path $P_{i,j}$ in $G$, 
    which has to also be the underlying path of the fastest temporal path from $v_i$ to $v_j$, and using \cref{obs:travel-delays-in-trees,lemma:unique-fastest-path-allDelays} calculate travel delays at all inner vertices.
    We save those values in a matrix $T$, of size $n \times n \times n$, and mark vertices of the path  $P_{i,j}$ as visited.
    Matrix $T$ stores at the position $(k,j,\ell)$ the value  $\tau_{v_k}^{v_{j}, v_\ell}$, where $v_j,v_\ell \in N(v_k)$. All other values of $T$ are set to \textsc{Null}.
    We repeat the above procedure from vertex $v_i$, for vertices $v_{j'}$ that are not marked as visited yet.
    When we have exhausted the $i$-th row of $D$,
    \ie vertex $v_i$ now reaches every vertex of $G$,
    we continue and repeat the procedure for all other vertices.
    If at any point we get two different numerical values for the same entry of $T$ we stop and return the negative answer.
    The above calculation is performed in polynomial time, as for every vertex $v_i$ we inspect the whole graph once.
    \begin{claim}[$\star$]\label{lemma:matrixT-travelDelays-tree}
        Matrix $T$ consists of travel delays for all vertices of degree at least $2$ in $G$.
    \end{claim}
    Given the matrix $T$, we can easily determine a labeling $\lambda$ that satisfies $D$.
    We start by fixing the label of an arbitrary edge as $a$, where $a \in [\Delta]$.
    Knowing the label of one edge, and all waiting times in $T$, we uniquely determine the labels of all other edges.
    More specifically, if we know that $\lambda(v_i v_j) = a$, then for all $v_k \in N(v_i)$ (resp.~$v_{k'} \in N(v_j)$)
    the value $\lambda(v_iv_k) \equiv a + \tau_{v_i}^{v_j,v_k} \pmod \Delta $ (resp.~$\lambda(v_j v_{k'}) \equiv a + \tau_{v_j}^{v_i,v_{k'}} \pmod \Delta $).
    Since there are $\Delta$ options to fix the first label, there are $\Delta$ different labelings satisfying $D$.
    %Note, w.l.o.g. we can start determining the labeling $\lambda$ by setting $\lambda(v_1v_2) = 1$.
    It is not hard to see, that also the calculation of the labeling $\lambda$ takes polynomial time.
    %as we have to traverse the graph exactly once, to successfully fix the labeling. 
    Therefore, all together the whole algorithm is performed in polynomial time.
\end{proof}

@@

In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Note that one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$.%, \ie $F = E(G) \setminus E(T)$.

@@

To overcome the above mentioned difficulties, we propose the following high-level strategy.
\begin{itemize}
    \item We identify a small number of ``interesting vertices'' between which we guess representative fastest paths.
    \item We show that we can determine most labels along each path up to an additive constant. We show that the number of labels we cannot determine along such a path is small. 
    \item For each additive constant and each non-determined label along such a path we introduce a variable for an Integer Linear Program (ILP). 
    \item We use constraints in the ILP to ensure that alternative connections (not using the guessed fastest path) between vertex pairs are not too fast.
    \item We compute a solution to the ILP and construct a labeling from that solution. Since the ILP has few variables and not too many constraints, this computation can be done efficiently using Lenstra's Theorem~\cite{Lenstra1983Integer}.
\end{itemize}

For the remainder of this section, let $D$ be the input matrix of \deltaExact\, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
%With $F$ we denote a minimum feedback edge set of $G$, and with $k$ the feedback edge number of $G$.
We are now ready to present our FPT algorithm. For an easier readability we split the description and analysis of the algorithm in five subsections.
We start with a preprocessing procedure for $G$, where we define a set of interesting vertices which then allows us to guess the desired structures.
Next we introduce some extra properties of our problem, that we then use to create ILP instances and their constraints.
At the end we present how to solve all instances and produce the desired labeling $\lambda$ of $G$, if possible.

@@

%%%%% HM: the two sentences below seem trivial to me.
%Note that we can easily solve \deltaExact\ by guessing the structure of temporal paths among all pairs of vertices.
%But, this would create $f(n,m)$ guesses (where $f(n,m)$ is an exponential function in $|V|=n$ and $|E| = m$), and the running time of the algorithm would be too big.
We approach the problem using some special properties, which allow us to create $f(k)$ different guesses for certain fastest temporal paths (where $f$ will be a double exponential function in the feedback edge number $k$).
We solve each guess by creating an ILP instance with $O(k^3)$ variables and $O(n^c)$ constraints, for some $c \in \mathbb{N}$.
If the ILP instance is a \textsc{Yes}-instance then we will show that also $D$ is a \textsc{Yes}-instance of \deltaExact.
Moreover, from a positive solution of the ILP instance we then construct the desired labeling $\lambda$.
If all constructed ILP instances are \textsc{No}-instances, then we will show that $D$ is a \textsc{No}-instance of \deltaExact.
Due to Lenstra's Theorem~\cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables.
Therefore, we can solve each of our ILP instances in FPT time with respect to $k$.
\end{comment}
%

\subparagraph{Preprocessing of the input.}
From $G$ we create $G'$ by
iteratively removing vertices of degree one. With $Z$ we denote the set of all removed vertices.
%by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
%Note that $F$ is also a minimum feedback edge set of $G$.
%Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, that gives rise to a minimum feedback edge set $F$ (denote $k = |F|$), 
and
let $V_1 \subset V(G')$ be the set of leaves in $T$, $V_2 \subset V(G')$ the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with 
$
U = V_1 \cup V_2 \cup V_3
$
the set of \emph{vertices of interest}. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie 
$
U^* = \{v \in V(G') \setminus U  :  u \in U, v \in N(u)\}.
$
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, therefore $|U^*| \in O(k^2)$.
Note that determining all of the above sets takes linear time.

\begin{comment}
From the construction it follows that $Z$ consists of disjoint trees $T_1, T_2, \dots$.
For a tree $T_i$ we denote with $u_i$ the vertex in $G'$ that is a neighbor of a vertex in $T_i$, 
and call it a \emph{clip vertex of the tree $T_i$}.
It follows that there can be many different trees $T_i$ that are incident to the same clip vertex $u_i \in V(G')$,
but each tree $T_i$ is incident to exactly one clip vertex $u_i \in V(G')$.
Since $u_i$ is the only vertex connecting all of the trees $T_i$ incident to it,
from now on we assume that a tree $T_{u_i}$ in $Z$ is a union of trees on vertices from $V(G) \setminus V(G')$,
that are clipped at the same vertex $u_i \in V(G')$.
For each of the trees $T_{u_i}$ in $Z$, we select one vertex $r_i$, that is a neighbor of the clip vertex $u_i$,
and call it \emph{a representative vertex of the tree $T_{u_i}$}.
We now define as $Z^*$ the set of representatives $r_i$ of trees $T_i \in Z$,
where the clip vertex $v_i$ of $T_i$ is a vertex of interest, \ie
$
Z^* = \{ r_i  :  r_i \in T_i, \text{ where $T_i \in Z$, the clip vertex $u_i$ of $T_i$ is in $U$, and $r_i u_i \in E(G)$}\}.
$
Since there are $O(k)$ vertices of interest, we get that $|Z^*| \in O(k)$.
Note that determining sets $U$, $U^*$, and $Z^*$ takes linear time. 
\end{comment}
\subparagraph{Guessing necessary structures. \label{sec:FPT-guessing}}
Once the sets $U, U^*$ %and $Z^*$ 
are determined, we are ready to start guessing  the necessary structures.
Note that whenever we say that we guess the fastest temporal path between two vertices, we mean that we guess the underlying path of a representative fastest temporal path between those two vertices.
In the case when we want to guess the fastest path from $u$, that (somehow) passes through the vertex $x$, and goes to $v$ via an edge (or a unique path), we write it as a fastest path of form $u \leadsto x \rightarrow v$.
%Similarly it holds if we want to include more vertices that have to be visited by the path. 
%If there is an edge (\ie a unique fastest path) among two vertices, we denote it by $\rightarrow$, if the fastest path among two vertices is not uniquely determined, we denote it by $\leadsto$.
We guess $11$ different structures, we present here just a few of them, for all the others we refer the reader to the Appendix.
\begin{enumerate}[G-1.]
    \item \label{FPT-guessFTPamongU}
    The fastest temporal paths among all pairs of vertices of $U$.
    For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
    Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path, respecting the value from $D$.
    %If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
    Repeating this procedure for all pairs $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ variations of the fastest temporal paths among all vertices in $U$.

    %\item \label{FPT-guessFTPamongZstar}
    %The fastest temporal paths among all vertices in $Z^*$, which by similar arguing as for vertices in $U$, gives us $O(k^{k^3})$ guesses.

    %\item \label{FPT-guessFTPamongUstar}
    %The fastest temporal paths among all vertices in $U^*$. This gives us $O(k^{k^6})$ guesses.
